Solutions to the potential for unintended conservation consequences of MPAs

### 2021-10-09 documents/03_marlin_paper.Rmd
OK, looking good. Knits with all base figure versions. back in the morning

---
title: "Solutions to the potential for unintended conservation consequences of MPAs"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: references.bib
params:
  run_name: ["v0.5"]
---

@@

@@

@@

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. The modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
increasingly being proposed specifically to benefit these by-catch
populations while minimizing losses or where possible providing benefits
to fisheries [citation].

@@

what we found was a series of neccesary but not sufficient conditions

make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.

@@

We focused first on the range of impacts that MPAs can have on mature
biomass across the range of variables explored. To illustrate, we pulled
out two example simulations, one where MPAs provide relatively large
benefits across all three species, and a second where MPAs actually
produced substantial losses in some species. In the "best case"
scenario, each of the three species shares a common habitat
distribution, and so protection of the central habitat of one species
results in protection for all species. In the "worst case" scenario,
each of the species have markedly different centroids. An MPA designed
primarily around the tuna and marlin populations then ends up displacing
effort into the shark habitat, producing benefits for the first two
species but driving the shark population down.

@@

the point is not once answer

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run, the results were highly variable, with MPAs capable of
producing substantial conservation gains to substantial conservation
losses and everything in between. We do not focus into the individual
effects of particular variables as we found them to be highly dependent
on the exact question asked and the set of data included, and since we
cannot ascribe any actual likelihood to any individual state of nature
simulated here.

## Case Study

For this subset of simulations then, along with the various "rule of
thumb" MPA design strategies we ran an optimization routine (see section
XX) across a range of objective function weightings, in the manner of
@rassweiler2014. This allowed us to assess 1) the degree to which
unintended conservation consequences of MPAs can be reduced by optimized
design and 2) the magnitude of tradeoffs between conservation outcomes
and fishery revenues under both rule-of-thumb and optimized MPA
networks.

key results: optimization helps (Nobel prize please). Can reduce percent
of species with conservation loss down to 25% pretty quickly if you
prioritize conservation. Rule of thumb designs like targeting fishing
can get there with perfect information and when considering all animals,
but can be pretty far from optimal with imperfect information or when
only considering a subset of animals in the planning process (that's
what all the different points represent). Some rules of thumb are more
sensitive to these decision than others. E.g. rate can get you optimal,
but goes off the rails pretty quickly with bad information.

There is also a pretty clear breakpoint in results between basically
trying for conservation or trying for economics, and not a lot of wiggle
room in between (the two general bands of lines). xx check to what
extent this depends on the way you scale the two objective function
components xx.

```{r, fig.cap="Percent of species with some conservation loss as a function of MPA size. Points are draws from rule-of-thumb scenarios, lines are results of optimization routine across a gradient of conservation weighting."}

@@

Key results: Even with optimization, you can basically either be in a
world with 50% or more of species having some level of unintended
conservation consequence, but no economic loss / some economic gain, or
you can be in a world with lower unintended consequences, but economic
losses. But, even rule of thumb approaches can get you to some places
with relatively little unintended consequences at relatively minimal
economic loss.

```{r, fig.cap="Tradeoff between unitendedn consequences and fisheries revenue. Filled polygons in background represent rule-of-thumb results, with redder colors indicating higher density of points. Points are results of optimization, with color indicating conservation weighting in optimization. Black diagonal line shows 1:1 relationship"}

@@

Key results: Presence of and magnitude of unintended consequences pretty
sensitive to design descisions. On a macro scale though far less so. In
terms of change in total SSB/SSB0 (metric of biodiversity, i.e. maxed
out when all species are unfished) and change in revenue, there is
basically a pretty simple linear tradeoff between biodiversity and
revenue, with a slope less than negative 1, meaning that 1% increase in
biodiversity comes at a greater than 1% loss in fishery revenue. This
stacks up pretty nicely with @rassweiler2014., where value of
optimization was low when overfishing of valuable species isn't
prevalent, as is the case here. So, design and optimization can help a
lot with the more "micro" level outcomes (e.g. making sure that
biodiversity gaines aren't coming from massive gains in one species to
detriment of other), but on a macro level pretty insensitive in this
case study.

```{r, fig.cap="Tradeoff frontier between revenue and biodiversity xx across different rule odf thumb design strategies. Points indicate role of thumb results, lines are results of optimized networks across a gradient of conservation weighting."}

@@

motivated by but not limited to pelagic ecosystems

@@

>>>>

---
title: "Solutions to the potential for unintended conservation consequences of MPAs"
subtitle: Conditions for success and unintended consequences in pelagic MPA design
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: references.bib
params:
  run_name: ["v1.1"]
---

@@

sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name)) %>% 
  bind_rows(.id = "critter") %>% 
  pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>% 
  mutate(common_name = stringr::str_to_title(common_name))

@@

c

@@

in the Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. The modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
increasingly being proposed specifically to benefit these by-catch
populations while minimizing losses or where possible providing benefits
to fisheries [citation].

@@

Our simulation experiments allow us to examine the wide range of effects
that MPAs can have on groups of species depending on combinations of
ecological, economic, and design parameters. The goal of this exercise
is not to assign any probability to the potential impacts of MPAs; we
have not provided any explicit probabilites to the alternative states
modeled here, and so it may be that a state that was only modeled on a
few occasions in our simulations is in fact highly common in the real
world.

What these experiments are designed to do is illustrate the range of
outcomes that are possible from MPAs under the types of conditions
evaluated here, and evaluate which drivers are potentially the most
influential in affecting particular outcomes, such as unintended
conservation consequences. Our case study of a WCPO-esque system is
intended then to reduce the parameter space enough that we can examine
the relative tradeoffs and value of optimization for economic and
conservation objectives in a pelagic system.

One clear caveat across all of these results is that we have made two
very strong assumptions around the behavior of the fishing fleet. The
first is that we assume that effort is displaced, not removed, by the
creation of MPAs. This is a widely debated question around MPAs, and the
real impacts of MPAs on fishing effort will clearly be setting
dependent. For example, large-scale pelagic fishing fleets already have
high rates of vessel movement built into their economic models, and so
may be more likely to move rather than exit the fishery in response to
MPAs. In contrast, smaller scale coastal fishing operations may simply
exit the fishery if their local fishing grounds are placed in MPAs.
There is clear evidence of effort displacement in many fisheries though
(citation), and moreover a motivation for this study was to evaluate the
potential unintended consequences of MPAs from effort displacement on
target and bycatch species. If MPAs do not produce increases in fishing
effort outside their borders, then the conservation effects of MPAs have
to be positive for all species concerned.

The second is that we do not account for evolution of total effort. Our
assumption is that while fleets can make economic decisions based around
where to allocate their effort in space, the total pool of effort they
exert is fixed (we would note that the underling `marlin` model does
have capacity for more realistic fleet models such as open-access
dynamics). This is clearly an incorrect assumption over the long term,
but allows us to focus on the impacts of the other drivers considered
here. A more through consideration of the impacts of fleet model
structure on MPA outcomes is generally needed, and is critical for any
tactical efforts at MPA modeling.

<!-- what we found was a series of neccesary but not sufficient conditions  -->

<!-- make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.  -->

@@

We focused first on the range of impacts that MPAs can have on mature
biomass across the range of variables explored. To illustrate, we pulled
out two example simulations, one where MPAs provided relatively large
benefits across all three species, and a second where MPAs actually
produced substantial losses in some species. In the "best case"
scenario, each of the three species shares a common habitat
distribution, and so protection of the central habitat of one species
results in protection for all species. In the "worst case" scenario,
each of the species have . An MPA designed primarily around the tuna and
marlin populations then ends up displacing effort into the shark
habitat, producing benefits for the first two species but driving the
shark population down XX (Fig.\@ref(fig:fig-info)).

@@

```{r}

critter_results <- compact_experiment_results %>%
  filter(bau_biodiv > 0, prop_mpa > 0) %>% 
  group_by(state_id, placement_id) %>%
  mutate(real_sigma_centroid = sd(centroid)) %>%
  mutate(delta_biodiv = pmin(1,biodiv - bau_biodiv),
         loss = biodiv < bau_biodiv,
         real_loss = biodiv < (0.9 * bau_biodiv),
         rec_form = as.factor(rec_form)) %>% 
  mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
         rec_form = fct_recode(rec_form, "Global DD, Rec. By Habitat" = "0", "Local DD, local Recruitment" = "1", "Local DD, Disperse Recruits" = "2", "Disperse Larvae, Poset-Settlement DD" = "3"))

@@

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run, the results were highly variable, with MPAs capable of
producing substantial conservation gains to substantial conservation
losses and everything in between. Across our simulations, gains were
much more common than losses, with losses of any magnitude occurring in
`r scales::percent(mean(critter_results$loss))` of simulations with
MPAs, but losses of 10% or greater of SSB in only
`r scales::percent(mean(critter_results$real_loss))` of simulations.
However we stress again that we can assign no probabilities to any of
these outcomes in the real world; this simply tells us that under the
range of conditions, selected to represent plausible states of nature,
there are many more ways for MPAs to provide a conservation effect
greater than or equal to 0 than there are for them to provide a net loss
for some species.

The question of interest then are which of the considered drivers appear
to be the most influential in affecting the magnitude and direction of
conservation effects of MPAs. Fishing mortality is the most obvious
answer; if fishing mortality is zero in the absence of an MPA then
clearly a no-take MPA will have no effect. As such we do not focus on
that metric here, focusing instead on other key drivers identified by
our simulation. Four visually prominent factors were MPA size, species,
design strategy, and degree of habitat difference. The degree to which
MPAs of a given size impact spawning biomass depends on a host of
factors. The potential for conservation losses are intuitively convex in
MPA size: effects of any kind are low when MPA size is low, and
conservation losses must be zero when MPA size is 100% (which this graph
doesnot quite reach), with peak losses in our simulations occuring
between MPA sizes of XX and XX. Life history plays a clear role our
stylized with shark and marlin populations being much more likely to
experience net conservation losses as a function of MPA placement than
the more resilient tunas. High degree of habitat differences can lead to
both to the potential for large conservation gains or losses depending
on whether a more isolated species is located primarily inside or
outside of MPAs (Fig.\@ref(fig:fig-exp-biodiv)).

```{r fig-exp-biodiv}

fig_exp_biodiv <- compact_experiment_results %>%
  filter(prop_mpa < 1) %>%
  # filter(f_v_m > 1.5) %>%
  group_by(state_id, placement_id) %>%
  ggplot(aes(
    prop_mpa,
    pmin(1,biodiv - bau_biodiv),
    color =  sigma_habitat,
    group = interaction(placement_id, state_id)
  )) +
  geom_line(alpha = 0.5) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  facet_grid(critter ~ placement_strategy) +
  scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
  scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
  theme(legend.position = "top")

# 
# compact_experiment_results %>%
#   filter(prop_mpa < 1) %>%
#   # filter(f_v_m > 1.5) %>%
#   group_by(state_id, placement_id) %>%
#   ggplot(aes(
#     prop_mpa,
#     econ - bau_econ,
#     color =  f_v_m,
#     group = interaction(placement_id, state_id)
#   )) +
#   geom_line(alpha = 0.5) +
#   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
#   facet_grid(critter ~ placement_strategy, scales = "free_y") +
#   scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
#   scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
#   scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
#   theme(legend.position = "top")

fig_exp_biodiv
```

We can somewhat quantitatively evaluate the relative impact of these
drivers using a regression, given that by construction variation in each
of evaluated drivers are independent in our simulations. We fit a
binomial regression to the presence or absence of of a net conservation
loss per species as a function of the simulated experimental variables.
The factors that according to this model most increase the probability
of a net conservation loss occurring include habitat differences, a
placement strategy of avoiding fishing areas, and the life history of
the species (shark or marlin relative to tuna). The factors that most
decrease the probability of a conservation loss from an MPA are MPA
size, placement strategy based on local depletion (xx might remove this
strategy as unrealistic xx), and somewhat interestingly fishing
mortality rate (Fig.\@ref(fig:fig-loss-drivers)-A).

However, this result is somewhat misleading (xx so should we just
remove?xx). Fig.\@ref(fig:fig-exp-biodiv) shows the clearly bimodal
nature of for example habitat differences. This analysis helps inform us
that habitat differences are a necessary but not sufficient condition
for MPA leakage induced conservation losses. The regression based
results require assuming that each of the effects are independent and
linear in nature. In fact each of these factors can interact in complex
and non-linear ways that may not be apparent visually or through this
type of regression analysis. To address this, we ran a random forest
using the `ranger` package in R (citaitons) and extracted the impurity
corrected importance scores for each of the variables. While this does
not tell us the expected marginal effect of individual drivers, it does
tell us which variables are critical if our goal is to accurately
predict the potential for unintended consequences or magnitude of
conservation gains in MPAs. Note here that variables that intuitively
should have an impact but have no visually or statistically (in a linear
model) clear impact on the probability of experiencing a conservation
loss such as the dispersal distances of adults or larvae have importance
scores comparable in magnitude to habitat differences when considered in
the context of nonlinear effects and interactions with other variables
(Fig.\@ref(fig:fig-loss-drivers)-B).

```{r fig-loss-drivers}

# loss_reg <-   rstanarm::stan_glm(
#   loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
#   family = binomial(),
#   data = critter_results %>% filter(between(prop_mpa,.2,.4)),
#   cores = 4,
#   chains = 4
# )

loss_reg <-   glm(
  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
  family = binomial(),
  data = critter_results %>% filter(between(prop_mpa,.2,.4))
)

loss_reg_coefs <- broom::tidy(loss_reg) %>%
  mutate(lower = estimate - 1.96 * std.error,
         upper = estimate + 1.96 * std.error) %>%
  mutate(
    term = str_replace(term, "placement_strategy", "Placement Strategy: "),
    term = str_replace(term, "rec_form", "DD Timing: "),
    term = str_replace(term, "spatial_allocation", "Fleet Allocation: "),
    term = str_replace(term, "hyper", "Hyperallometry present"),
    term = str_replace(term, "prop_mpa", "MPA size"),
    term = str_replace(term, "sigma_habitat", "Habitat differences"),
     term = str_remove_all(term, "critter")
  )

loss_coef_plot <- loss_reg_coefs %>%
  ggplot() +
  geom_hline(aes(yintercept = 0)) +
  geom_pointrange(aes(
    forcats::fct_reorder(term, estimate),
    estimate,
    ymin = lower,
    ymax = upper
  )) +
  scale_y_continuous(name = "Log-odds Ratio Effect") +
  scale_x_discrete(name = '') +
  coord_flip()

tmp_critter <-
  critter_results %>% filter(between(prop_mpa, .2, .4)) %>%
  select(-habitat) %>% 
  mutate(fctr_loss = factor(loss))
# 
# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
#   recipes::juice()

loss_rf <-   ranger::ranger(
  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
  data = tmp_critter,
  importance = "impurity"
)

loss_importance <- data.frame(importance = loss_rf$variable.importance %>% sort()) %>% 
  rownames_to_column(var = "term")

loss_importance_plot <- loss_importance %>% 
  ggplot(aes(forcats::fct_reorder(term, importance), importance)) + 
  geom_col() +
  scale_y_continuous(name = "Variable Importance") +
  scale_x_discrete(name = '') +
  coord_flip()

fig_loss_drivers <- loss_coef_plot + loss_importance_plot

fig_loss_drivers
```

```{r fig-delta-drivers, eval = FALSE}

delta_reg <-   rstanarm::stan_glm(
  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + hyper,
  data = critter_results %>% filter(between(prop_mpa,.2,.4)),
  cores = 4,
  chains = 4
)

tmp_critter <-
  critter_results %>% filter(between(prop_mpa, 0.2, 0.4)) %>%
  select(-habitat) %>% 
  mutate(fctr_loss = factor(loss))
# 
# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
#   recipes::juice()

delta_rf <-   ranger::ranger(
  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
  data = tmp_critter,
  importance = "impurity"
)

delta_rf$variable.importance %>% sort()

plot(delta_reg)

@@

## WCPO Case Study

The generalized experiments help us understand the potential impacts of
the evaluated drivers, but also highlight the challenge in making
definitive statements about the effect of a given driver on the outcomes
of pelagic(?) MPAs, given the complex ways in which all of these factors
interact, even within a relatively simple model such as this one. In
order to help illustrate how addressing the kinds of spatial dynamics
and technical interactions included in our model can affect and benefit
marine spatial planning, we then ran a case study intended to loosely
reflect the dynamics of WCPO tuna, billfish, and shark fisheries and
populations.

This case study allows us to focus in on important questions around the
relative performance of "rule of thumb" MPA design strategies relative
to optimized networks. We ran a spatial optimization routine (see
section XX) across a range of objective function weightings, in the
manner of @rassweiler2014. This allowed us to assess 1) the potential
risk of unintended conservation consequences from MPAs in this WCPO
bycatch study and the 2) assess value of optimization in this WCPO case
study in achieving economic and conservation objectives.

Given the set of parameters simulated here, we find little intrinsic
risk for large-scale unintended losses in total spawning biomass caused
by MPA effort displacement in the WCPO case study. Each of the design
scenarios produced at least some small loss in biomass for one or more
species at some MPA size. However, these losses were generally small and
isolated to a subset of species within a small MPA size range, with the
exception of the "Avoid_fishing" MPA placement strategy that resulted in
net losses for most species at large MPA sizes (though simply minimal
impact at smaller sizes). All design strategies except "Avoid_fishing"
and "Rate" resulted in relatively rapid declines in profits for the
three main tuna fisheries (skipjack, yellowfin, and bigeye). Both
"Avoid_fishing" and "Rate" mainlined profits close to business as usual
up to a large MPA size, but did so at the expense of conservation
benefits (Fig.\@ref(fig:fig-cs-critters)).

```{r fig-cs-critters, fig.cap = "Simulated trajectories of changes in profits and spawning stock biomass (SSB) for selected species caught in the WCPO. Each column represents a different MPA design strategy, see Table.XX for definitions. Optimized refers to an MPA network optimized to maximize total SSB across all species."}

opt_comparison <- opt_obj %>%
  filter(alpha == max(alpha)) %>%
  mutate(placement_error = 0,
         prop_mpa = p_protected / 100,
         placement_strategy = "Optimized") %>%
  select(critter, placement_error, delta_biodiv, delta_econ, prop_mpa,placement_strategy)

capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}

 experiment_obj %>% 
   select(colnames(opt_comparison)) %>% 
  bind_rows(opt_comparison) %>% 
   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
   left_join(sci_to_com, by = "critter") %>% 
  filter(placement_error == min(placement_error)) %>% 
   rename("Change in SSB" = delta_biodiv, "Change in Profits" = delta_econ) %>% 
  pivot_longer(c("Change in SSB", "Change in Profits"), names_to = "metric", values_to = "value") %>% 
  ggplot(aes(prop_mpa, value, color = common_name)) + 
   scale_color_discrete(name = '') +
   geom_hline(aes(yintercept = 0), linetype = 2)+
  geom_line(show.legend = TRUE) + 
   scale_y_continuous(name = '') +
   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
   theme(strip.text.x = element_text(size = 8), 
         legend.position = "top", 
         legend.text = element_text(size = 8))

Viewed as totals, we found a small parameter space under the "Area" and
"Avoid_fishing" design strategies where MPAs were capable of causing a
net decrease in total spawning stock biomass across all evaluated
species. However, the efficiency frontier indicates that these points
are not Pareto efficient, and MPA design configurations exist that
achieve the same levels of economic performance while producing a net
gain in spawning biomass. Optimal MPA design was able to slow the
tradeoff between total profits and bycatch spawning biomass, but for any
conservation weighting any increases in total biomass of bycatch species
came at a decrease in total profits. However we also found limited need
for optimized networks, with the "Depletion" design scenario essentially
following the Pareto frontier (Fig.\@ref(fig:fig-cs-totals)).

```{r fig-cs-totals, fig.cap = "Efficiency frontiers for changes in total profits and total bycatch spawning stock biomass (SSB). Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function. Red points show outcomes of 'rule of thumb' MPA design strategies. "}
 total_opt_obj %>%
  group_by(alpha) %>%
  filter(obj_fun == max(obj_fun)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(total_bycatch_ssb_change, delta_econ, color = alpha), size = 1.5) +
  geom_point(
    data = total_experiment_obj,
    aes(total_bycatch_ssb_change, delta_econ),
    size = 2,
    alpha = 0.1,
    fill = "red",
    shape = 21
  ) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_abline(intercept = 0, slope = -1) +
  facet_wrap(~placement_strategy, scales = "free_y", labeller = labeller(placement_strategy = capitalize)) + 
  scale_color_viridis_c(name = "Conservation Weighting in Optimization",
                        guide = guide_colorbar(frame.colour = "black", ticks.colour = "black", barwidth = unit("15", "lines"))) +
  scale_x_continuous(name = "Change in Total Bycatch SSB") +
  scale_y_continuous(name = "Change in Total Profits ") +
  theme(axis.text.x = element_text(size = 8),
        legend.position = "top")

```{r, eval = FALSE}

 total_opt_obj %>%
  group_by(alpha) %>%
  filter(obj_fun == max(obj_fun)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(delta_biodiv, delta_econ, color = alpha), size = 2) +
  geom_point(
    data = total_experiment_obj,
    aes(delta_biodiv, delta_econ),
    size = 2,
    alpha = 0.1,
    fill = "red",
    shape = 21
  ) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_abline(intercept = 0, slope = -1) +
  facet_wrap( ~ placement_strategy) +
  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
  scale_y_continuous(name = "Change in Total Profits Relative to BAU ") +
  theme(axis.text.x = element_text(size = 8),
        legend.position = "top")

@@

MPAs can be a powerful tool for providing conservation benefits across
large swaths of species and space. Under the appropriate circumstances
they can also provide benefits to food security and economic outcomes.
However, while extremely simple in concept, predicting the likely
effects of MPAs across groups of species and fishing fleets becomes
increasingly complicated as the systems being studied deviate from
assumptions of homogenously distributed species with perfectly selective
fleets. In this paper we used a new bio-economic model to demonstrate
how a range of variables often omitted in MPA modeling can act to
influence the outcomes of MPAs on conservation and economic objectives.
We demonstrate that unintended conservation losses are possible, but
under a range of specific conditions, and under most simulated cases did
not occur. MPAs were capable of causing small losses in biomass for the
WCPO case study, but only for a subset of design strategies and species,
most of which could be overcome by optimized design.

Our results indicate that as suggested by theory (citation) and
empirical evidence (citation), MPAs can indeed have unintended
consequences for conservation if they result in effort displacement
rather than attrition. However, our model found these to be a minority
of the simulations tested. This does in no way suggests that leakage may
not a problem for specific MPA. But, it also indicates that across the
range of plausible parameter combinations that we explored it was much
more common for MPAs to provide net conservation benefits even if effort
is displaced. This suggests then that we should of course not dismiss
the conservation benefits of MPAs in the presence of effort
displacement, but also that users must be aware of the potential for
unintended consequences that can indeed occur. While many different
factors interact in complex ways to affect the potential for unintended
conservation consequences, the most obvious is the degree of difference
in the habitats utilized by the species within an ecosystem in which an
MPA is to be placed. As a starting point, it is critical that
communities engaged in MSP processes create a localized assessment of
the degree of habitat overlap across potentially affected species to
assess the potential risk of unintended consequences.

The MPA literature is highly dominated by modeling studies (citation).
What then is the added value of the model and results presented here?
Our model contains many features missing from many general MPA models,
including size structure, seasonal movement, larval and adult habitats
and dispersal, and critically technical interactions xx clearer term for
this xx across species. A reader might reasonably ask about a range of
other factors not included, such as time-varying parameters and species
interactions. One could of course continue to add features to the model
until it reaches the complexity of an end-to-end model such as Atlantis
(citation). While such models have tremendous value, our goal here is to
present a model that can help provide closer to tactical advice for
users without the data or capacity for such high resolution models. One
justification for simpler MPA models is often computational
tractability: tracking the full age structure of multiple species and
the economic outcomes of multiple fleets dynamically across a high
resolution spatial grid can quickly become computationally expensive.
Faster implementations completely written in languages such as C++
however become inaccessible to many users. Our model seeks to fill a
middle ground here, with the core operations run in C++ while providing
a relatively user-friendly front end in the open-source and commonly
used R programming language. This allows users to ask the sorts of
questions addressed in this paper that require a more realistic model,
without requiring inordinate amounts of data, programming expertise, or
computational time.

Caveats

Note that lack of importance of density dependence timing is conditional
on age structure, and matters more if you're doing surplus production
modeling (cite letter)

Mention hyperallometry and distinction to what that study did

Perfect information

Fleet model

@@

We do not focus into the individual effects of particular variables as
we found them to be highly dependent on the exact question asked and the
set of data included, and since we cannot ascribe any actual likelihood
to any individual state of nature simulated here.

```{r}
